Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2010-12-31 rebuttal.tex

\section{Response to Reviewer \#1}
Reviewer 1 makes specific recommendations for improving the
presentation of the paper, many of which we have heeded.  Please refer to the change
list for details.

\paragraph{About Presentation:} The reviewer suggests that we can reduce the description
of the ALM algorithm in Section II-D. We agree that ALM has been extensively studied in the optimization
literature. We do not claim any novelty here. However, we believe a complete description of the algorithm
here is important for the TPAMI audience: There are
many algorithms for solving the associated optimization problem, with drastically different complexity
and speed. Typical vision researchers might not be so familiar with some of the recent
developments in these algorithms. In particular some of the more effective improvements to ALM are fairly
recent and scattered in different literatures. In fact, many of the questions raised by reviewer \#2
ask for even more detailed explanation about the algorithms. So, we believe that a complete but brief
description of the ALM-based algorithm is important for the audience of TPAMI to implement and
test the ideas efficiently and correctly.

We find the reviewer's other suggestions on presentation useful and have made changes accordingly.

\paragraph{About Positioning and Novelty:} As we have tried to make clear, the
proposed method does not aim to solve face recognition for all scenarios! It is
not designed for the case where the training is uncontrolled and limited, and
hence it may not have any advantages over other face recognition
techniques/systems specifically designed for such scenarios.  The proposed
system is an attempt to build a system that is not only accurate and robust
enough to compete with other biometric measures but also discreet and
non-intrusive when in use. There are many practical scenarios where the
proposed setting is very relevant, such as access control, where well
controlled training images can (and should) be obtained in advance but no well
controlled test images are needed or available. Particularly, the system should
be able to handle typical daily changes in facial appearance, such as occlusion
by glasses, hats, or facial hair and variations in lighting and moderate
expression. The possibility of achieving fully automated (at recognition time),
touch-free identification is arguably the main advantage of face recognition
over other biometric measures such as fingerprint and iris recognition.  To the
best of our knowledge even for this admittedly restricted scenario, there is no
practical face recognition system whose performance can meet the standards of
most access control and security applications. Existing systems have found
success  mostly in applications that are less demanding in performance, such as
online image search or photo album organization.

@@

Thanks to the reviewer's comments, we have modified the introduction to make this point more clear, especially clarifying the specific scenario where our method can claim novelty.

\paragraph{About Summary Recommendation:} We agree with the reviewer that there is a need for a careful evaluation of our system and comparison to other techniques or systems for face recognition. However, because most face recognition systems are not designed for the scenario considered by this paper, a fair and meaningful comparison is difficult. Nevertheless, to clarify both the strengths and limitations of our approach, we have augmented the experimental section with comparisons to Local Binary Patterns (LBP), in addition to the comparisons already performed. LBP is one of the most popular tools for face verification, and is used in many commercial face recognition systems.

We compare both methods with the same settings on two datasets that are representative of controlled and uncontrolled training: the Multi-PIE and the FERET
datasets.  The results indicate both suboptimal performance of LBP for applications where many controlled training images (under different illuminations) are available, and conversely, suboptimal performance of our algorithm on the FERET dataset where the training is neither controlled nor sufficient.

\paragraph{Detailed Concerns or Comments by Reviewer 1:} \begin{enumerate}
\item Reviewer 1 is correct that use of our system is largely restricted to
users who are cooperative (in the sense of non-adversarial). Again, the main
point of our paper is that we can make identification more reliable and
convenient, at the relatively small cost of doing a better job of gathering
data at enrollment time.  Ideally, face recognition will take place
automatically as a user walks through a door or works in front a desktop.

@@

In addition, if the accuracy and reliability of a face recognition system can
truly get close to that of fingerprint and iris, it can serve as a
complementary measure in a multi-modal system and improve the overall security
for high-stakes applications. To the best of our knowledge, no existing face
recognition system, with or without cooperative training, comes close to meet
the standards of high-stake security applications.

In response to the reviewer's question "is occlusion really a problem?", we have found that the use of
the L1-norm as a robust error function improves performance even when the test
image does not contain occlusions.  Any boundaries where the intensity changes
in a misaligned image can result in a large, localized error component.  While
unoccluded faces are mostly smooth, sharp boundaries still occur where there
are cast shadows, especially at depth discontinuities such as the nostrils, the
sides of the nose, and partially open mouths.

@@

\item Rigorous justification of the choice of illuminations is a subject of
continuing research.  Our current choice is motivated empirically by the
experiments described in the paper.

\item To our knowledge, we are the first to use projectors for illumination in
face recognition applications.  Section 1A of the previous submission already
cites several other applications that use projectors as illumination sources.
The requested references propose a technique for improving the SNR of projector-based
illumination systems for certain classes of objects.  We have added
discussion of this to the paper.

\item  It is true that one or both stages of the algorithm could be substituted
with other techniques.  That said, the registration and recognition stages are
coherent in the sense that they are motivated by same modeling assumptions
(illumination model, sparse occlusion, etc.), they use the same robust penalty
on image error (the L1-norm), and they use the same numerical optimization
techniques (ALM).  While conceptual simplicity is a secondary concern in
algorithm design, starting with a minimalist foundation benefits both
practitioners as well as researchers.  Other researchers have already published
more complex extensions to our work.

\item  {\bf There is no extensive computation involved in
    Step 13 of Algorithm 1. In fact, for each candidate
    subject $i\in {k_1, \ldots, k_s}$ for recognition, Step
    13 simply uses the transformation $\tau_i$ obtained
    from the iterative alignment stage (Step 2-9) to
    re-sample each original training
    image of that subject, so that this new set of training
    samples is aligned with the test sample. Nevertheless,
    it worth mentioning that, since we perform alignment
    separately for each class in the gallery, the
    computation for the alignment stage (Step 2-9) is
    linear in the size of the gallery. However, thanks to
    its separability, the alignment stage can be easily
    parallelized.}

    The variation in alignment across classes is indeed
    small, but still changes the objective enough to affect
    recognition results. Aligning to a generic face (or a
    gallery based on ``aligned'' images of many people) is
    a good idea, and one we have considered as well.  If
    nothing else, it could be used to improve alignment
    initialization instead of using raw face detector
    output.  It may also be useful to use the roughly
    aligned test image to quickly eliminate some of the
    classes from further consideration.  Again, we have
    deliberately tried to keep the algorithm as minimalistic
    as possible for this publication.

@@

\item (iii) ``How often the solution gets drifted away into a wrong basin of
attraction. (If there is any way to know this.)''

In Section II E, we have conducted careful experiments to measure the region of attraction
for the proposed alignment method. We have studied the region of attraction of the proposed
method for both 2D and 3D pose variation, as well as how it can be improved with a proper
multi-scale implementation.

\item (iv) ``What is the criteria or threshold for convergence? If the threshold is
stringent, the algorithms are likely to get stuck in infinite loops, and if the
threshold is liberal, premature convergence would occur. How are such
thresholds picked and how universal are they? Does every new dataset need its
own threshold?''

{\bf For the $\ell_1$ minimizations we terminate the ALM
algorithm when the relative changes of primal variables in
two consecutive iterations become small, as suggested by
[27]. That is,
\begin{displaymath}
\frac{\|\x_{k+1}-\x_k\|_2}{\|\x_k\|_2} < \epsilon \quad
\textup{and} \quad \frac{\|\e_{k+1}-\e_k\|_2}{\|\e_k\|_2} <
\epsilon,
\end{displaymath}
where $\epsilon>0$ is the tolerance. With the suggested
criteria, the l1-minimizations are guaranteed to converge
eventually, although convergence may be very slow if they
are chosen poorly. For the alignment
iteration, we terminate when the change of objective value
(\ie $\|\e\|_1$) in two consecutive iterations becomes
small.} We use the same set of convergence thresholds for
all the data sets we ran against.

@@

\end{enumerate}

\section{Response to Reviewer \#3}
Reviewer 3 makes the following requests:

\begin{enumerate}
\item ``I would have liked to see the experiment for the case S=1, without
impostors in the dataset, to judge the effect of the ``Sparse
Concentration Index''.

{\bf We have added experiments to the paper demonstrating
the utility of the sparse representation(SR)-based
recognition stage for classification.  In general, we see a
few percentage points of improvement in recognition rate
when using the representation residual obtained from the
SR-based recognition stage (with $S=10$) over the
recognition based purely on alignment error (with $S$=1).
While this effect is desirable, as discussed in the paper,
it is worthy emphasizing that the SR-based recognition also
enables us to compute SCI, which has demonstrated its power
for impostor rejection.}

@@

In addition, the sparse representation techniques are not limited to frontal 2D images.
They can be easily applied to 2D images from multiple poses or to establish alignment
between a 2D image and a 3D face model (if available). These are all good directions
for future research.

We have modified Section I.A and the Conclusion to address these issues.

\end{enumerate}

\end{document}
 a user's face than a single low-resolution depth image, especially if
recognition is to be performed on 2D input images.

In addition, the sparse representation techniques are not limited to frontal 2D images.
They can be

### 2010-12-31 faces_system.tex
Changes from Zihan

We briefly review existing techniques for recognition, image registration, and
handling of illumination variation.  Our system is based purely on 2D
techniques.  This fact immediately distinguishes our approach from systems
that either require a 3D data as an input, or attempt to estimate a 3D model
from 2D input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
techniques can achieve better robustness to pose variation given
a sufficiently accurate 3D model, for access control applications where only 
moderate misalignment and pose are present, the proposed method will be more
than sufficient.   Note that according to the principle of photometric stereo, images of faces
under varying illuminations already contain 3D shape-related information,\footnote{In principle,
one can recover the 3D shape of the face from multiple illuminations using photometric stereo.} 
and this information can be leveraged by 2D algorithms for alignment and recognition 
even if shape is not reconstructed explicitly. In addition, the proposed sparsity-based robust
alignment and recognition techniques are not limited to 2D data and can be easily extended
to 3D data if necessary.

@@

\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector. We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
\begin{equation}
\y\circ \tau + J \Delta \tau = A_i \x + \e.
\end{equation}
Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
Jacobian of $\y \circ \tau$ with respect to the transformation
parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
above equation is underdetermined if we allow the registration
error $\e$ to be arbitrary. Near the correct alignment we
expect the aligned testing image to differ from $A_i \x$ only
for the minority of the pixels corrupted by occlusions. Thus,
we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
\begin{equation}
\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
\label{eqn:L1-align}
\end{equation}
This is different from the popular choice that
minimizes the $\ell^2$-norm of the registration error:
\begin{equation}
\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
\label{eqn:L2-align}
\end{equation}
which is also equivalent to finding the deformation step
$\Delta  \tau$ by solving the least-square problem:
$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
\|_2$. Empirically, we find that if there is only small noise
between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
$\eqref{eqn:L2-align}$ have similar performance.  However, if
there are occlusions in $\y_0$, sequential
$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
better than sequential $\ell^2$-minimization
\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
example.

@@

CMU Multi-PIE provides the most extensive test set among public
datasets. This database contains images of 337 subjects across
simultaneous variation in pose, expression, and illumination.
Of these 337 subjects, we use all of the 249 subjects present
in Session 1 as the training set. The remaining 88 subjects are
treated as ``impostors'', or invalid images. For each of the
249 training subjects, we include frontal images of 7 frontal
illuminations,\footnote{They are illuminations
$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
directional illumination, we subtract the ambient-illuminated
image 0.} taken with neutral expression. As suggested by the
work of \cite{Georghiades2001-PAMI}, these extreme frontal
illuminations would be sufficient to linearly represent other
frontal illuminations, as will also be corroborated by the next
experiment on our own dataset. For the test set, we use all 20
illuminations from Sessions 2-4, which were recorded at
distinct times over a period of several months. The dataset is
challenging due to the large number of subjects, and due to
natural variation in subject appearance over time. Table
\ref{tab:MultiPIE-recognition} shows the result of our
algorithm on each of the 3 testing sessions. Our algorithm
achieves recognition rates above $90\%$ for all three sessions,
with input {\em directly} obtained from the Viola and Jones'
face detector; there was no manual intervention for the test images.
To demonstrate that the SCI-based recognition step is indeed beneficial
even when there are no impostors, we include results for recognition
based only on the alignment error residuals, shown in row 6.

\begin{table}
\caption{Recognition rates on large-scale Multi-PIE database.}
\vspace{-.2in}
\centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & Session 2 & Session 3 & Session 4  \\
\hline
\hline
LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
\hline
NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
\hline
NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
\hline
LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
\hline
\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
\hline
{Algorithm 1 w/o SCI} & 90.7\% & 89.6\% & 87.5\% \\
\hline
{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
\hline
{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
\hline
\end{tabular}
\label{tab:MultiPIE-recognition} }
\end{table}
\begin{figure}
\centering
{\small
\begin{tabular}{cccccc}
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{-.1in} \\
(a) & (b) & (c) & (d) & (e) & (f) \vspace{-.1in}
\end{tabular}
}
\caption{{\bf Representative failures from Multi-PIE}. {\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2. Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
\label{fig:failed-examples}
\vspace{-.5in}
\end{figure}

@@

We test the algorithms' ability to reject invalid images of the
88 subjects not appearing in the training database. As
mentioned before, the \emph{sparsity concentration index} (SCI)
is used as the outlier rejection rule. Given the sparse
representation $\x$ of a test image with respect to $K$
training classes, the SCI measures how concentrated the
coefficients are on a single class in the dataset and is
defined as in \cite{Wright2009-PAMI}:
\begin{displaymath}
\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
\|\x\|_1 - 1}{K - 1} \in [0,1] .
\end{displaymath}
It is easy to see that if $\textup{SCI}(\x) = 1$, the test
image is represented using images from one single subject
class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
evenly over all classes. Thus, we can choose a threshold $t \in
[0,1]$ for the proposed method and accept a test image as valid
if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
invalid. We compare this classifier to classifiers based on
thresholding the error residuals of NN, NS and LDA.

@@

>>>>

We briefly review existing techniques for recognition, image registration, and
handling of illumination variation.  Our system is based purely on 2D
techniques.  This fact immediately distinguishes our approach from systems
that either require a 3D data as an input, or attempt to estimate a 3D model
from 2D input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While these
techniques can achieve better robustness to pose variation given
a sufficiently accurate 3D model, for access control applications where only 
moderate pose variation is present, the proposed method will be more
than sufficient.   Note that according to the principle of photometric stereo, images of faces
under varying illuminations already contain 3D shape-related information,\footnote{In principle,
one can recover the 3D shape of the face from multiple illuminations using photometric stereo.} 
and this information can be leveraged by 2D algorithms for alignment and recognition 
even if shape is not reconstructed explicitly. In addition, the proposed sparsity-based robust
alignment and recognition techniques are not limited to 2D data and can be easily extended
to 3D data if necessary.

@@

\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector. We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
\begin{equation}
\y\circ \tau + J \Delta \tau = A_i \x + \e.
\end{equation}
Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
Jacobian of $\y \circ \tau$ with respect to the transformation
parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
above equation is underdetermined if we allow the registration
error $\e$ to be arbitrary. At the correct alignment we
expect the test image to differ from $A_i \x$ only
for the minority of the pixels corrupted by occlusions. Thus,
we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
\begin{equation}
\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
\label{eqn:L1-align}
\end{equation}
This is different from the popular choice that
minimizes the $\ell^2$-norm of the registration error:
\begin{equation}
\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
\label{eqn:L2-align}
\end{equation}
which is also equivalent to finding the deformation step
$\Delta  \tau$ by solving the least-square problem:
$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
\|_2$. Empirically, we find that if there is only small noise
between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
$\eqref{eqn:L2-align}$ have similar performance.  However, if
there are occlusions in $\y_0$, sequential
$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
better than sequential $\ell^2$-minimization
\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
example.

@@

CMU Multi-PIE provides the most extensive test set among public
datasets. This database contains images of 337 subjects across
simultaneous variation in pose, expression, and illumination.
Of these 337 subjects, we use all of the 249 subjects present
in Session 1 as the training set. The remaining 88 subjects are
treated as ``impostors'', or invalid images. For each of the
249 training subjects, we include frontal images of 7 frontal
illuminations,\footnote{They are illuminations
$\{0,1,7,13,14,16,18\}$ of \cite{Gross2008-FGR}. For each
directional illumination, we subtract the ambient-illuminated
image 0.} taken with neutral expression. As suggested by the
work of \cite{Georghiades2001-PAMI}, these extreme frontal
illuminations would be sufficient to linearly represent other
frontal illuminations, as will also be corroborated by the next
experiment on our own dataset. For the test set, we use all 20
illuminations from Sessions 2-4, which were recorded at
distinct times over a period of several months. The dataset is
challenging due to the large number of subjects, and due to
natural variation in subject appearance over time. Table
\ref{tab:MultiPIE-recognition} shows the result of our
algorithm on each of the 3 testing sessions. Our algorithm
achieves recognition rates above $90\%$ for all three sessions,
with input {\em directly} obtained from the Viola and Jones'
face detector; there was no manual intervention for the test images.
To demonstrate that the sparse representation based recognition step is indeed beneficial
even when there are no impostors, we include results for recognition
based only on the alignment error residuals(i.e. S=1), shown in row 6.

\begin{table}
\caption{Recognition rates on large-scale Multi-PIE database.}
\vspace{-.2in}
\centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & Session 2 & Session 3 & Session 4  \\
\hline
\hline
LDA$_d$ (LDA$_m$) & 5.1 (49.4)\%  & 5.9 (44.3)\% & 4.3 (47.9)\%  \\
\hline
NN$_d$ (NN$_m$)  & 26.4 (67.3)\% & 24.7 (66.2)\% & 21.9 (62.8)\%  \\
\hline
NS$_d$ (NS$_m$) &  30.8 (77.6)\% & 29.4 (74.3)\% & 24.6 (73.4)\% \\
\hline
LBP$_d$ (LBP$_m$) & 39.9 (93.3)\% & 38.1 (91.2)\% & 33.9 (92.9)\% \\
\hline
\cite{Yang2010-CVPR} & {\bf 95.2}\% & 93.4\% & 95.1\% \\
\hline
{Algorithm 1, $S=1$} & 90.7\% & 89.6\% & 87.5\% \\
\hline
{Algorithm 1} & 93.9\% & 93.8\% & 92.3\% \\
\hline
{Algorithm 1 with improved window} & 95.0\% & {\bf 96.3}\% & {\bf 97.3}\% \\
\hline
\end{tabular}
\label{tab:MultiPIE-recognition} }
\end{table}
\begin{figure}
\centering
{\small
\begin{tabular}{cccccc}
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_01_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_01_01_051_08.png} \\
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/079_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/111_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/196_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/130_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/163_02_01_051_08.png}  &
\includegraphics[width=0.9in,clip=true]{figures_pami/multipie_failed/175_02_01_051_08.png} \vspace{-.1in} \\
(a) & (b) & (c) & (d) & (e) & (f) \vspace{-.1in}
\end{tabular}
}
\caption{{\bf Representative failures from Multi-PIE}. {\bf Top:} training from Session 1; {\bf Bottom:} test images from Session 2. Due to changes in hair, glasses, beard, or pose, our alignment fails on these subjects regardless of test image illumination.}
\label{fig:failed-examples}
\vspace{-.5in}
\end{figure}

@@

@@

We test the algorithms' ability to reject invalid images of the
88 subjects not appearing in the training database. As
mentioned before, the \emph{sparsity concentration index} (SCI)
is used as the outlier rejection rule. Given the sparse
representation $\x$ of a test image with respect to $K$
training classes, the SCI measures how concentrated the
coefficients are on a single class in the dataset and is
defined as in \cite{Wright2009-PAMI}:
\begin{displaymath}
\textup{SCI}(\x) \doteq \frac{K \cdot \max_i \|\delta_i(\x)\|_1 /
\|\x\|_1 - 1}{K - 1} \in [0,1] .
\end{displaymath}
It is easy to see that if $\textup{SCI}(\x) = 1$, the test
image is represented using images from one single subject
class; if $\textup{SCI}(\x) = 0$, the coefficients are spread
evenly over all classes. Thus, we can choose a threshold $t \in
[0,1]$ for the proposed method and accept a test image as valid
if $\textup{SCI}(\x) \geq t$, and otherwise reject it as
invalid. We compare this classifier to classifiers based on
thresholding the error residuals of NN, NS, LDA, and LBP.
